Capstone Movielens Report

### 2025-02-18 reports/capstone-movielens-report.Rmd
Accounting for Date & Genre effects (in progress).

@@

>>>>

library(dslabs)
library(tidyverse)

@@

We use the following models to account for the `date` effect:

$$
Y_{i,j} = \mu + \alpha_i + \beta_j + f(d_{i,j}) + \varepsilon_{i,j}
$$

### Accounting for Genre effect
\
As mentioned in [Section 23.7: Exercises](https://rafalab.dfci.harvard.edu/dsbook-part-2/highdim/regularization.html#exercises) of the _Chapter "23 Regularization" of the Course Textbook_ the `Movielens` dataset also has a genres column. This column includes every genre that applies to the movie (some movies fall under several genres)[@IDS2_23-7].

The plot below shows strong evidence of a genre effect (for illustrative purposes, the plot shows only categories with more than 20, 000 ratings).

```{r}
# Preparing data for plotting:
train_set_grp <- train_set |> 
  mutate(genre_categories = as.factor(genres)) |>
  group_by(genre_categories) |>
  summarize(n = n(), avg = mean(rating), se = sd(rating)/sqrt(n())) |>
  filter(n > 20000) |> 
  mutate(genres = reorder(genre_categories, avg)) |>
  select(genres, avg, se, n)

dim(train_set_grp)
print(train_set_grp |> sort_by.data.frame(~ avg))

# Creating plot:
train_set_grp |> 
  ggplot(aes(x = genres, y = avg, ymin = avg - 2*se, ymax = avg + 2*se)) + 
  geom_point() +
  geom_errorbar() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
Below are worst and best ratings categories:
```{r}
sprintf("The worst ratings are for the genre category: %s",
        train_set_grp$genres[which.min(train_set_grp$genres)])

sprintf("The best ratings are for the genre category: %s",
        train_set_grp$genres[which.max(train_set_grp$genres)])
```

If we define $g_{i,j}$ as the genre for user's $i$ rating of movie $j$, we can use the following models to account for the `genre` effect:

$$
Y_{i,j} = \mu + \alpha_i + \beta_j + f(d_{i,j}) + \sum_{k=1}^K x_{i,j} \beta_k + \varepsilon_{i,j}
$$

with $x^k_{i,j} = 1$ if $g_{i,j}$ is genre $k$.
